---
title: "Analysis of movies- IMDB dataset"
output: html_document
slug: proj3
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest) # to scrape wikipedia page
```

# Analysis of movies- IMDB dataset

We will look at a subset sample of movies, taken from the [Kaggle IMDB 5000 movie dataset](https://www.kaggle.com/carolzhangdc/imdb-5000-movie-dataset)

  
```{r,load_movies, warning=FALSE, message=FALSE}

movies <-read_csv(here::here("data","movies.csv")) 
glimpse(movies)

```

Besides the obvious variables of `title`, `genre`, `director`, `year`, and `duration`, the rest of the variables are as follows:

- `gross` : The gross earnings in the US box office, not adjusted for inflation
- `budget`: The movie's budget 
- `cast_facebook_likes`: the number of facebook likes cast memebrs received
- `votes`: the number of people who voted for (or rated) the movie in IMDB 
- `reviews`: the number of reviews for that movie
- `rating`: IMDB average rating 

- Are there any missing values (NAs)? Are all entries distinct or are there duplicate entries?
```{r, movies}
skimr::skim(movies)
# the number of double entries based on movies
length(subset(movies$title,duplicated(movies$title)))

movies_unique<- movies %>% distinct(title, year, .keep_all = TRUE)
length(subset(movies_unique$title,duplicated(movies_unique$title)))
# View(movies_unique)
```


- Produce a table with the count of movies by genre, ranked in descending order
```{r}
library(tidytext)
library(janeaustenr)
library(dplyr)
#https://dplyr.tidyverse.org/reference/count.html
movies_unique %>% 
    count(genre,wt = NULL, sort = TRUE, name = NULL) #%>% # group by genre
```


- Produce a table with the average gross earning and budget (`gross` and `budget`) by genre. Calculate a variable `return_on_budget` which shows how many $ did a movie make at the box office for each $ of its budget. Ranked genres by this `return_on_budget` in descending order
```{r}

tab1<-movies_unique %>% 
    group_by(genre) %>% # group by genre
    summarise(gross = mean(gross, na.rm = TRUE), 
              budget = mean(budget, na.rm = TRUE),
              return_on_budget=gross/budget)

arrange(tab1, desc(return_on_budget))
  
```


- Produce a table that shows the top 15 directors who have created the highest gross revenue in the box office. Don't just show the total gross amount, but also the mean, median, and standard deviation per director.
```{r}
movies_unique %>% 
  group_by(director) %>%
  summarise(  tot=sum(gross,na.rm = TRUE),
              avg = mean(gross, na.rm = TRUE), 
              median = median(gross, na.rm = TRUE),
              std=sd(gross))%>%
  slice_max(order_by =(tot), n=15)
  
```


- Finally, ratings. Produce a table that describes how ratings are distributed by genre. We don't want just the mean, but also, min, max, median, SD and some kind of a histogram or density graph that visually shows how ratings are distributed. 
```{r}
dat1<-movies_unique %>% 
  group_by(genre) %>%
  summarise(  avg=mean(rating,na.rm = TRUE),
              min = min(rating, na.rm = TRUE), 
              max = max(rating, na.rm = TRUE),
              median = median(rating, na.rm = TRUE),
              std=sd(rating,na.rm = TRUE))

dat1

movies_unique %>% 
ggplot( aes(x=genre, y=rating))+
  geom_boxplot()+
  geom_jitter(width=0.15)+
  theme(axis.text.x = element_text(angle = 90))+
   theme_bw()+
  labs(
    title = "Rating by genre",
    subtitle = "Documentary perform well",
    x = "genre",
    y = NULL)

```

  - Examine the relationship between `gross` and `cast_facebook_likes`. Produce a scatterplot and write one sentence discussing whether the number of facebook likes that the cast has received is likely to be a good predictor of how much money a movie will make at the box office. What variable are you going to map to the Y- and X- axes?
  
  
```{r, gross_on_fblikes}
movies_unique %>% 
  ggplot(aes(x = gross, y = cast_facebook_likes, gross)) +
  geom_point(size=1, shape=1) +
  theme_bw()+
  labs(
    title = "Relationship between gross profit and cast_facebook_like",
    subtitle = "No relationship",
    x = "Gross earnings (US$)",
    y = "Facebook likes"
  )

```
**There is no relationship between the two variables**

  - Examine the relationship between `gross` and `budget`. Produce a scatterplot and write one sentence discussing whether budget is likely to be a good predictor of how much money a movie will make at the box office.

```{r, gross_on_budget}

movies_unique %>% 
  ggplot(aes(x = gross, y = budget, gross)) +
  geom_jitter(width = 10000000, height = 10000000)+
  labs(
    title = "Relationship between budget and cast_facebook_like",
    subtitle = "Positive relationship",
    x = "Gross earnings (US$)",
    y = "Budget"
  )
```
**There is a positive relationship between the two variables** 

  - Examine the relationship between `gross` and `rating`. Produce a scatterplot, faceted by `genre` and discuss whether IMDB ratings are likely to be a good predictor of how much money a movie will make at the box office. Is there anything strange in this dataset?

```{r, gross_on_rating}
movies_unique %>% 
  ggplot(aes(x = gross, y = rating, gross)) +
  geom_point(size=1, shape=1) +
  theme_bw()+
  labs(
    title = "Relationship between gross profit and cast_facebook_like",
    subtitle = "No relationship",
    x = "Gross earnings (US$)",
    y = "Facebook likes"
  )+facet_wrap(~genre, scales = "free")
```

