---
categories:
- ""
- ""
title: "Epidemic Simulation"
output: html_document
slug: proj1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)

# default figure size
knitr::opts_chunk$set(
  fig.width=15, 
  fig.height=5,
  fig.align = "center"
)
```

Simple code to simulate COVID-19 infection in the society.  


Eliminate the exponential notation of the integer.
```{r}
options(scipen = 10)
```

Number of citizens
```{r}
N_population <- 1000
```


Number of initial positive cases
```{r}
N_init <- 2
```


Death rate and infection rate
```{r}
death_rate <- 0.002
infection_rate <- 0.3
```


Number of days to recover
```{r}
n_recover <- 10
```

Constants (State of health)
```{r}
healthy <- 0
infected <- 1
dead <- 2
```


Number of people encountered by the infected person in one day.(A number randomly selected from 0 to N_encountered)
```{r}
N_encountered <- 20
```


Immunity coefficient: A probability of being infected for immune person is the infection rate multiplied by the immunity coefficient.
```{r}
immunity_coef <- 0.3
```


```{r}
sim <- function(death_rate=0.002, infection_rate=0.3, n_recover=10, N_encountered=20){
  # Construct one r data frame storing id, state, duration, immunity.
  id <- seq(1,N_population,1)
  state <- rep(0,N_population)
  duration <- rep(0,N_population)
  immunity <- rep(1,N_population)
  data <- data.frame(id, state, duration, immunity)
  # Initialization
  id_init <- ceiling(runif(N_init,min=0,max=N_population))
  data[id_init,]$state <- infected
  # vectors storing number of infected person and deaths
  N_Infected <- c()
  N_New_Infected <- c()
  N_Death <- c()
  N_New_Death <- c()
  N_Immunity <- c()
  N_New_Immunity <- c()
  for(i in 1:300){
    # calculate number of infected person and deaths
    n_infected <- sum(data$state == infected)
    n_death <- sum(data$state == dead)
    n_immunity <- sum(data$immunity == immunity_coef)
    
    # update duration
    data[data$state == infected,3] = data[data$state == infected,3] + 1
    
    # update immunity
    # data[data$state == infected,4] = immunity_coef
    
    # update death
    death_id <- runif(n_infected,0,1) < death_rate
    n_new_death <- sum(death_id)
    data[data$state == infected,][death_id,2] <- dead
    #print(death_id)
    #data[death_id,2] <- dead
    
    # infect new citizens
    N_meet <- floor(runif(1,0,N_encountered*n_infected))
    meet_id <- unique(ceiling(runif(N_meet,min=0,max=N_population)))
    # remove the dead
    meet_id <- meet_id[data[meet_id,2] != dead]
    # remove the currently infected
    meet_id <- meet_id[data[meet_id,2] != infected]
    # people with immunity
    meet_id_1 <- meet_id[data[meet_id,4] == immunity_coef]
    infected_id_1 <- meet_id[runif(length(meet_id_1),0,1) < infection_rate * immunity_coef]
    # people without immunity
    meet_id_2 <- meet_id[data[meet_id,4] == 1]
    infected_id_2 <- meet_id[runif(length(meet_id_2),0,1) < infection_rate]
    # total
    infected_id <- c(infected_id_1,infected_id_2)
    n_new_infected <- length(infected_id)
    
    # update new infected citizens
    data[infected_id,2] <- infected
    
    # patients recovery
    n_new_immunity <- nrow(data[data$duration >= n_recover,])
    data[data$duration >= n_recover,2] <- healthy
    data[data$duration >= n_recover,4] <- immunity_coef
    data[data$duration >= n_recover,3] <- 0
    
    # display information
    # print(sprintf(fmt = "Day: %3d; Total Infections: %6d; Toatl Deaths: %6d; Total Recovered:%6d; New Infections: %5d; New Deaths: %5d; New Recovered: %5d", i, n_infected, n_death, n_immunity, n_new_infected, n_new_death, n_new_immunity))
    
    # Stopping Criterion
    # if(n_infected == 0 | n_death == N_population){
    #   print("Stopping criterion reached.")
    #   break
    # }
    
    # Append number of infected, death, and immunity
    N_Infected <- append(N_Infected,n_infected)
    N_Death <- append(N_Death,n_death)
    N_Immunity <- append(N_Immunity,n_immunity)
    
    N_New_Infected <- append(N_New_Infected, n_new_infected)
    N_New_Death <- append(N_New_Death, n_new_death)
    N_New_Immunity <- append(N_New_Immunity, n_new_immunity)
    
  }
  days <- seq(1,length(N_Infected))
  d <- data.frame(days, N_Infected, N_Death, N_Immunity, N_New_Infected, N_New_Death, N_New_Immunity)
  return(d)
}
```

The first simulation: Death rate is 0.2%, infection rate is 30%,   
number of days to recover is 10, number of people contacted in one day is 20  
```{r}
d1 <- sim(death_rate=0.002, infection_rate=0.3, n_recover=10, N_encountered=20)
```

The second simulation: Death rate is 3%, infection rate is 30%,   
number of days to recover is 10, number of people contacted in one day is 20  
```{r}
d2 <- sim(death_rate=0.03, infection_rate=0.3, n_recover=10, N_encountered=20)
```

The third simulation: Death rate is 0.2%, infection rate is 5%,   
number of days to recover is 10, number of people contacted in one day is 20  
```{r}
d3 <- sim(death_rate=0.002, infection_rate=0.05, n_recover=10, N_encountered=20)
```

The fourth simulation: Death rate is 0.2%, infection rate is 30%,   
number of days to recover is 20, number of people contacted in one day is 20  
```{r}
d4 <- sim(death_rate=0.002, infection_rate=0.3, n_recover=20, N_encountered=20)
```

The fifth simulation: Death rate is 0.2%, infection rate is 30%,   
number of days to recover is 10, number of people contacted in one day is 2  
```{r}
d5 <- sim(death_rate=0.002, infection_rate=0.3, n_recover=10, N_encountered=2)
```

Take a glance at the first several lines of the result data frame
```{r}
d1[1:30,]
```

A function used to combine many plots
```{r}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  plots <- c(list(...), plotlist)
  numPlots = length(plots)
  if (is.null(layout)) {
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}
```


Plot the results of five simulations.
```{r}
p1 <- ggplot(d1, aes(x=days, y=N_Infected)) + 
  geom_line(aes(y=N_Infected, color = "Normal")) + 
  geom_line(aes(y=d2$N_Infected, color = "High Death Rate")) +
  geom_line(aes(y=d3$N_Infected, color = "Low Infection Rate")) +
  geom_line(aes(y=d4$N_Infected, color = "Longer Recovery Days")) +
  geom_line(aes(y=d5$N_Infected, color = "Low Contact Numbers")) +
  scale_color_manual(name = "Conditions", 
                     values=c("Normal" = "#FF6600", 
                              "High Death Rate" = "#999900", 
                              "Low Infection Rate" = "#33CC33",
                              "Longer Recovery Days" = "#3399FF", 
                              "Low Contact Numbers" = "#FF33CC")) +
  ggtitle("Total Infections")

p2 <- ggplot(d1, aes(x=days, y=N_Death)) + 
  geom_line(aes(y=N_Death, color = "Normal")) + 
  geom_line(aes(y=d2$N_Death, color = "High Death Rate")) +
  geom_line(aes(y=d3$N_Death, color = "Low Infection Rate")) +
  geom_line(aes(y=d4$N_Death, color = "Longer Recovery Days")) +
  geom_line(aes(y=d5$N_Death, color = "Low Contact Numbers")) +
  scale_color_manual(name = "Conditions", 
                     values=c("Normal" = "#FF6600", 
                              "High Death Rate" = "#999900", 
                              "Low Infection Rate" = "#33CC33",
                              "Longer Recovery Days" = "#3399FF", 
                              "Low Contact Numbers" = "#FF33CC")) +
  ggtitle("Total Death")


p3 <- ggplot(d1, aes(x=days, y=N_Immunity)) + 
  geom_line(aes(y=N_Immunity, color = "Normal")) + 
  geom_line(aes(y=d2$N_Immunity, color = "High Death Rate")) +
  geom_line(aes(y=d3$N_Immunity, color = "Low Infection Rate")) +
  geom_line(aes(y=d4$N_Immunity, color = "Longer Recovery Days")) +
  geom_line(aes(y=d5$N_Immunity, color = "Low Contact Numbers")) +
  scale_color_manual(name = "Conditions", 
                     values=c("Normal" = "#FF6600", 
                              "High Death Rate" = "#999900", 
                              "Low Infection Rate" = "#33CC33",
                              "Longer Recovery Days" = "#3399FF", 
                              "Low Contact Numbers" = "#FF33CC")) +
  ggtitle("Total Recovery")


p4 <- ggplot(d1, aes(x=days, y=N_New_Infected)) + 
  geom_line(aes(y=N_New_Infected, color = "Normal")) + 
  geom_line(aes(y=d2$N_New_Infected, color = "High Death Rate")) +
  geom_line(aes(y=d3$N_New_Infected, color = "Low Infection Rate")) +
  geom_line(aes(y=d4$N_New_Infected, color = "Longer Recovery Days")) +
  geom_line(aes(y=d5$N_New_Infected, color = "Low Contact Numbers")) +
  scale_color_manual(name = "Conditions", 
                     values=c("Normal" = "#FF6600", 
                              "High Death Rate" = "#999900", 
                              "Low Infection Rate" = "#33CC33",
                              "Longer Recovery Days" = "#3399FF", 
                              "Low Contact Numbers" = "#FF33CC")) +
  ggtitle("New Infections")

p5 <- ggplot(d1, aes(x=days, y=N_New_Death)) + 
  geom_line(aes(y=N_New_Death, color = "Normal")) + 
  geom_line(aes(y=d2$N_New_Death, color = "High Death Rate")) +
  geom_line(aes(y=d3$N_New_Death, color = "Low Infection Rate")) +
  geom_line(aes(y=d4$N_New_Death, color = "Longer Recovery Days")) +
  geom_line(aes(y=d5$N_New_Death, color = "Low Contact Numbers")) +
  scale_color_manual(name = "Conditions", 
                     values=c("Normal" = "#FF6600", 
                              "High Death Rate" = "#999900", 
                              "Low Infection Rate" = "#33CC33",
                              "Longer Recovery Days" = "#3399FF", 
                              "Low Contact Numbers" = "#FF33CC")) +
  ggtitle("New Death")

p6 <- ggplot(d1, aes(x=days, y=N_New_Immunity)) + 
  geom_line(aes(y=N_New_Immunity, color = "Normal")) + 
  geom_line(aes(y=d2$N_New_Immunity, color = "High Death Rate")) +
  geom_line(aes(y=d3$N_New_Immunity, color = "Low Infection Rate")) +
  geom_line(aes(y=d4$N_New_Immunity, color = "Longer Recovery Days")) +
  geom_line(aes(y=d5$N_New_Immunity, color = "Low Contact Numbers")) +
  scale_color_manual(name = "Conditions", 
                     values=c("Normal" = "#FF6600", 
                              "High Death Rate" = "#999900", 
                              "Low Infection Rate" = "#33CC33",
                              "Longer Recovery Days" = "#3399FF", 
                              "Low Contact Numbers" = "#FF33CC")) +
  ggtitle("New Recovery")

multiplot(p1, p4, p2, p5, p3, p6, cols=3)

```
