---
title: "GDP components over time and among countries"
output: html_document
slug: proj4
---


```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```

```{r load-libraries, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(GGally)
library(readxl)
library(here)
library(skimr)
library(janitor)
library(broom)
library(tidyquant)
library(infer)
library(openintro)
```

At the risk of oversimplifying things, the main components of gross domestic product, GDP are personal consumption (C), business investment (I), government spending (G) and net exports (exports - imports). You can read more about GDP and the different approaches in calculating at the [Wikipedia GDP page](https://en.wikipedia.org/wiki/Gross_domestic_product).

The GDP data we will look at is from the [United Nations' National Accounts Main Aggregates Database](https://unstats.un.org/unsd/snaama/Downloads), which contains estimates of total GDP and its components for all countries from 1970 to today. We will look at how GDP and its components have changed over time, and compare different countries and how much each component contributes to that country's GDP. The file we will work with is [GDP and its breakdown at constant 2010 prices in US Dollars](http://unstats.un.org/unsd/amaapi/api/file/6) and it has already been saved in the Data directory. Have a look at the Excel file to see how it is structured and organised

```{r read_GDP_data}

UN_GDP_data  <-
  read_excel(
    here::here("data", "Download-GDPconstant-USD-countries.xls"),
    # Excel filename
    sheet = "Download-GDPconstant-USD-countr",
    # Sheet name
    skip = 2
  ) # Number of rows to skip

```

The first thing you need to do is to tidy the data, as it is in wide format and you must make it into long, tidy format. Please express all figures in billions (divide values by `1e9`, or $10^9$), and you want to rename the indicators into something shorter.

```{r reshape_GDP_data}

tidy_GDP_data  <- UN_GDP_data %>%
  #change wider to longer by year
  pivot_longer(cols = c(4:51),
               names_to = "year",
               values_to = "value_e") %>%
  #into billions
  mutate(value = value_e / 1e9)

glimpse(tidy_GDP_data)

# Let us compare GDP components for these 3 countries
country_list <- c("United States", "India", "Germany")
```

First, can you produce this plot?

```{r}
#create list of needed components
component_list <-
  c(
    "Gross capital formation",
    "Exports of goods and services",
    "General government final consumption expenditure",
    "Household consumption expenditure (including Non-profit institutions serving households)",
    "Imports of goods and services"
  )

needed_GDP_data <- tidy_GDP_data %>%
  #get needed data filtered by country and components
  filter(Country %in% country_list) %>%
  filter(IndicatorName %in% component_list) %>%
  #turn `IndicatorName` from character to a factor variable
  mutate(indicator = factor(
    IndicatorName,
    levels = c(
      "Gross capital formation",
      "Exports of goods and services",
      "General government final consumption expenditure",
      "Household consumption expenditure (including Non-profit institutions serving households)",
      "Imports of goods and services"
    ),
    labels = c(
      "Gross capital formation",
      "Exports",
      "Government expenditure",
      "Household expenditure",
      "Imports"
    )
  )) %>%
  #change "year" into numerical
  mutate(year_n = as.numeric(year))


#make plot
needed_GDP_data %>%
  ggplot(aes(x = year_n, y = value, colour = indicator)) +
  geom_line(size = .8) +
  facet_wrap( ~ Country, nrow = 1) +
  theme_bw() +
  labs(
    title = "GDP components over time",
    subtitle = "In constant 2010 USD",
    x = "",
    y = "Billion US$",
    color = "Components of GDP"
  ) +
  scale_x_continuous(breaks = seq(1970, 2017, 10))
```

Secondly, recall that GDP is the sum of Household Expenditure (Consumption *C*), Gross Capital Formation (business investment *I*), Government Expenditure (G) and Net Exports (exports - imports). Even though there is an indicator `Gross Domestic Product (GDP)` in your dataframe, I would like you to calculate it given its components discussed above.

```{r}
#change longer to wider
needed_GDP_data <- needed_GDP_data[, c(2, 6, 7, 8)] %>%
  pivot_wider(names_from = "indicator", values_from = "value")
#calculate net exports
needed_GDP_data <- needed_GDP_data %>%
  mutate(net_exports = Exports - Imports)
#generate GDP
needed_GDP_data$GDP <- rowSums(needed_GDP_data[, c(3, 4, 5, 8)])

#calculate proportion for each component
needed_GDP_data <- needed_GDP_data %>%
  mutate(
    G_ratio = `Government expenditure` / GDP,
    C_ratio = `Household expenditure` / GDP,
    I_ratio = `Gross capital formation` / GDP,
    net_exports_ratio = net_exports / GDP
  )

```

> What is the % difference between what you calculated as GDP and the GDP figure included in the dataframe?

```{r}
#get the GDP figure included in the dataframe
gdp_origin <- tidy_GDP_data[, c(2, 3, 4, 6)] %>%
  filter(IndicatorName == "Gross Domestic Product (GDP)") %>%
  filter(Country %in% country_list) %>%
  mutate(year_n = as.numeric(year),
         compare_name = paste(Country, year_n))

#create a new variable to make it easier to join two data
needed_GDP_data <- needed_GDP_data %>%
  mutate(compare_name = paste(Country, year_n))

#join our gdp and origin gdp
mydata_compare <-
  left_join(gdp_origin, needed_GDP_data, by = "compare_name") %>%
  #calculate difference
  mutate(difference_gdp = (GDP - value) / value * 100)

#create a new dataset to see clearly
#only data needed
show_difference <- mydata_compare[, c(1, 3, 4, 14, 20)]
#change variable name
names(show_difference)[3:4] <- c("origin_GDP", "calculated_GDP")

show_difference
```

**We notice from dataset 'show_difference' that, for Germany and US, the GDP we calculated is always larger than GDP initially included.**

```{r}
#we try to reproduce the plot use 'needed_GDP_data'
needed_GDP_data_plot  <- needed_GDP_data %>%
  #change wider to longer by indicator, easier to plot
  pivot_longer(cols = c(10:13),
               names_to = "indicator",
               values_to = "proportion")

#change the name of legend
needed_GDP_data_plot$indicator <-
  factor(
    needed_GDP_data_plot$indicator,
    levels = c('G_ratio', 'C_ratio', 'I_ratio', "net_exports_ratio"),
    labels = c(
      "Government expenditure",
      "Household expenditure",
      "Gross capital formation",
      "Net exports"
    )
  )

#make plot, as we did above
needed_GDP_data_plot %>%
  ggplot(aes(x = year_n, y = proportion, color = indicator)) +
  geom_line(size = .8) +
  scale_color_manual(
    values = c(
      "Government expenditure" = "#F8766D",
      "Gross capital formation" = "#7CAE00",
      "Household expenditure" = "#00BFC4",
      "Net exports" = "#C77CFF"
    )
  ) +
  facet_wrap( ~ Country, nrow = 1) +
  theme_bw() +
  labs(
    title = "GDP and its breakdown at constant 2010 prices in US Dollars",
    x = "",
    y = "proportion",
    caption = "Source:United Nations, https://unstats.un.org/unsd/snaama/Downloads",
    color = ""
  ) +
  scale_x_continuous(breaks = seq(1970, 2017, 10)) +
  scale_y_continuous(labels = scales::percent)

```

> What is this last chart telling you? Can you explain in a couple of paragraphs the different dynamic among these three countries?

**For these three countries, household consumption is the largest element of expenditure across the economy, and we think it is also an important indicator of a country's economy landscape. As a country develops, the proportion of household expenditure, which is the expense of people's daily basic need, will go down, and other proportion will increase. This is pretty much like the Engel's law. For India, the proportion of household expenditure is the highest among three, which shows it is still a developing country. However, with the fast speed it developed during the late 20th century, this proportion went down and capital formation increased. A negative net exports shows that it imports a greater value than it exports, it has a trade deficit. For developed countries Germany and US, their graphs are somehow alike. However, Germany's net exports increased a lot since 21th century, which means Germany's goods are so competitive. We believe it is because reputation for quality precision engineering gives them a competitive advantage in selling goods to foreign countries, as well as the relatively weak euro.**

> If you want to, please change `country_list <- c("United States","India", "Germany")` to include your own country and compare it with any two other countries you like

