---
title: "Returns of financial stocks"
output: html_document
slug: proj3
---

```{r, setup, echo=FALSE}
knitr::opts_chunk$set(
  message = FALSE, 
  warning = FALSE, 
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
options(digits = 3)

# default figure size
knitr::opts_chunk$set(
  fig.width=6.75, 
  fig.height=6.75,
  fig.align = "center"
)
```


```{r load-libraries, warning=FALSE, message=FALSE, echo=FALSE}
library(tidyverse)  # Load ggplot2, dplyr, and all the other tidyverse packages
library(mosaic)
library(ggthemes)
library(lubridate)
library(fivethirtyeight)
library(here)
library(skimr)
library(janitor)
library(vroom)
library(tidyquant)
library(rvest) # to scrape wikipedia page
```

# Returns of financial stocks

```{r load_nyse_data, message=FALSE, warning=FALSE}
nyse <- read_csv(here::here("data","nyse.csv"))
```

Based on this dataset, create a table and a bar plot that shows the number of companies per sector, in descending order

```{r companies_per_sector}

tab<-nyse%>%
  count(sector, wt = NULL, sort = TRUE, name = NULL)
tab

tab%>%
  ggplot(aes(x = n, y=fct_reorder(sector,n))) +
  geom_col(fill='blue') +
  theme_bw()+
  labs(
    title = "Numbers of Companies per Sector",
    subtitle = "Finance dominates the pack",
    x = "count of companies",
    y = NULL
  )


```


```{r, tickers_from_wikipedia}

djia_url <- "https://en.wikipedia.org/wiki/Dow_Jones_Industrial_Average"


#get tables that exist on URL
tables <- djia_url %>% 
  read_html() %>% 
  html_nodes(css="table")


# parse HTML tables into a dataframe called djia. 
# Use purr::map() to create a list of all tables in URL
djia <- map(tables, . %>% 
               html_table(fill=TRUE)%>% 
               clean_names())


# constituents
table1 <- djia[[2]] %>% # the second table on the page contains the ticker symbols
  mutate(date_added = ymd(date_added),
         
         # if a stock is listed on NYSE, its symbol is, e.g., NYSE: MMM
         # We will get prices from yahoo finance which requires just the ticker
         
         # if symbol contains "NYSE*", the * being a wildcard
         # then we just drop the first 6 characters in that string
         ticker = ifelse(str_detect(symbol, "NYSE*"),
                          str_sub(symbol,7,11),
                          symbol)
         )

# we need a vector of strings with just the 30 tickers + SPY
tickers <- table1 %>% 
  select(ticker) %>% 
  pull() %>% # pull() gets them as a sting of characters
  c("SPY") # and lets us add SPY, the SP500 ETF

```


```{r get_price_data, message=FALSE, warning=FALSE, cache=TRUE}
# Notice the cache=TRUE argument in the chunk options. Because getting data is time consuming, # cache=TRUE means that once it downloads data, the chunk will not run again next time you knit your Rmd

myStocks <- tickers %>% 
  tq_get(get  = "stock.prices",
         from = "2000-01-01",
         to   = Sys.Date()) %>% # Sys.Date() returns today's price
  group_by(symbol) 

glimpse(myStocks) # examine the structure of the resulting data frame
```

```{r calculate_returns, message=FALSE, warning=FALSE, cache=TRUE}
#calculate daily returns
myStocks_returns_daily <- myStocks %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "daily", 
               type       = "log",
               col_rename = "daily_returns",
               cols = c(nested.col))  

#calculate monthly  returns
myStocks_returns_monthly <- myStocks %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "monthly", 
               type       = "arithmetic",
               col_rename = "monthly_returns",
               cols = c(nested.col)) 

#calculate yearly returns
myStocks_returns_annual <- myStocks %>%
  group_by(symbol) %>%
  tq_transmute(select     = adjusted, 
               mutate_fun = periodReturn, 
               period     = "yearly", 
               type       = "arithmetic",
               col_rename = "yearly_returns",
               cols = c(nested.col))
```

Create a table where you summarise monthly returns for each of the stocks and `SPY`; min, max, median, mean, SD.

```{r summarise_monthly_returns}

summarise_monthly_returns <- myStocks_returns_monthly %>% 
  #Consolidate data frame by symbol
  group_by(symbol) %>% 

  #Collapse data by calculating summary statistics  
  summarize(
    min = min(monthly_returns, na.rm = TRUE),
    median = quantile(monthly_returns, 0.5, na.rm = TRUE),
    max = max(monthly_returns, na.rm = TRUE),
    mean = mean(monthly_returns, na.rm = TRUE),
    sd = sd(monthly_returns, na.rm = TRUE))
summarise_monthly_returns

```


Plot a density plot, using `geom_density()`, for each of the stocks
```{r density_monthly_returns}

ggplot(myStocks_returns_monthly, aes(x = monthly_returns)) +
  geom_density(color = "black", fill = "transparent")+
    facet_wrap(~symbol, nrow = 4)

```

What can you infer from this plot? Which stock is the riskiest? The least risky? 

**The riskiest stock has the most spread out return performance - AAPL.**

**The least risky stock has the least spread out return performance - SPY.**

Finally, make a plot that shows the expected monthly return (mean) of a stock on the Y axis and the risk (standard deviation) in the X-axis. Please use `ggrepel::geom_text_repel()` to label each stock

```{r risk_return_plot}
summarize_monthly <- myStocks_returns_monthly %>% 
  group_by(symbol) %>% 
  summarise(aver_monthly_returns = mean(monthly_returns), 
            sd_monthly_returns = sd(monthly_returns))

#Compute graph
p <- ggplot(summarize_monthly, aes(x=sd_monthly_returns, y=aver_monthly_returns, label = symbol)) +
  geom_point(color = "red")

#Add aesthetics elements
library(ggrepel)
p1 <- p + geom_text_repel() + labs(title = "Mean-risk distribution", x="Standard Deviation", y="Mean")

p1
```

What can you infer from this plot? Are there any stocks which, while being riskier, do not have a higher expected return?

**CSCO + INTC. They have high risk, being at the higher end of SD measurement, whilst displaying suppressed returns.**
